Present bias in exploratory choice

### 2017-01-29 doc/cogsci2017/paper.tex
paper first draft - finished discussion

Under-exploration also seems to occur in the development of complex skills, such
as flying a plane or playing a sport CITE. In these settings, an ``emphasis
change'' training method in which the focus of practice is repeatedly changed
can lead to greater performance gains than unguided practice or more complex
training methods. Emphasis change works because it encourages people to
continually explore the performance space to find more effective cognitive and
motor routines CITE. Without these interventions, people often enter a ``local
maximum'' in which exploration decreases and performance plateaus CITE.

@@

Thirty-one participants completed the experiment, which was conducted over
Amazon Mechanical Turk using the psiTurk framework \citep{Gureckis2015a}.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.

@@

Forty participants completed the experiment, which was conducted over
Amazon Mechanical Turk using the psiTurk framework \citep{Gureckis2015a}.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.

@@

One explanation for our results is that people are in fact less affected by
present bias in the realm of exploratory choice than they are in other
inter-temporal choices.

>>>>

Under-exploration also seems to occur in the development of complex skills, such
as flying a plane or playing a sport \citep{Gopher1989}. In these settings, an ``emphasis
change'' training method in which the focus of practice is repeatedly changed
can lead to greater performance gains than unguided practice or more complex
training methods. Emphasis change works because it encourages people to
continually explore the performance space to find more effective cognitive and
motor routines. Without these interventions, people often enter a ``local
maximum'' in which exploration decreases and performance plateaus \citep{Yechiam2001}.

@@

Thirty-one participants completed the experiment, which was conducted over
Amazon Mechanical Turk (AMT) using the psiTurk framework \citep{Gureckis2015a}.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.

@@

Forty participants completed the experiment, which was conducted over
AMT using the psiTurk framework \citep{Gureckis2015a}.
Participants were paid \$5.00 for their participation, with a performance-based
bonus of up to \$3.00. All participants received the full \$3.00 bonus.

@@

One explanation for our results is that people are in fact less affected by
present bias in the realm of exploratory choice than they are in other
inter-temporal choices. This might be the case if exploratory choice is driven
less by an explicit present--future tradeoff and more by intrinsic ``exploration
bonuses'' \citep{kakade2002dopamine} or motivational states such as curiosity
\citep{Berlyne1966}. If exploration is driven by these mechanisms, then the
reward of exploration might be exploration itself rather than its later
consequences, breaking the intertemporal tradeoff between exploration and
exploitation.

However, given the early stage of the present work, we are not ready to make
strong claims about the role of present bias in exploratory choice. First, while
participants enjoyed the video more than the slider task, we need to perform a
follow-up study to ensure that this difference in enjoyment was strong enough to
create a bias towards immediate video-watching in the time scale used in our
experiments. To test this, we might use a simple ``concurrent schedule''
experiment in which two buttons produce video-watching time at for different
lengths (or with different probabilities) and at different delays. In this
setting, we would predict a bias towards the button that produced earlier
video-watching time, even if the video time was shorter or less frequent.
\citet{Navarick1998} used a similar paradigm to demonstrate present bias in
watching cartoons, though the bias might be reduced in participants on AMT due
to the greater availability of environmental distractors during non-video time.

We also may need to improve our decision-making task. As discussed, some
participants may have found the delayed condition confusing, and this problem
might be fixed through clearer instructions or interface. Additionally, the
future encounters with each machine may not have been sufficiently salient,
reducing the perceived need for exploration in both conditions.

While introduces difficulties not found in standard decision-making tasks, we
remain excited about the possibility of creating a paradigm for the use of
immediately-consumable rewards in dynamic decision-making tasks. Such a paradigm
could allow not just a test of whether present bias affects exploratory choice,
but a greater focus on a variety of motivational factors often neglected by
cognitive science.
